NumPy---Array Programming in Python

### 2020-01-11 body.tex
Minor edits

The {\em shape} of an array determines the number of elements along each axis,
and the number of axes is the array's dimensionality.
For example, a vector of numbers can be stored as a one dimensional array of
shape $N$, while color videos are four dimensional arrays of shape
$T \times M \times N \times 3$.

{\em Strides} are necessary to interpret computer memory, which stores elements
linearly, as $n$-dimensional arrays.
It describes the number of bytes to move forward in memory to jump from row to
row, column to column, and so forth.
Consider, for example, a 2-D array of floating point numbers with shape
$4 \times 3$, where each element occupies 8 bytes in memory.
To move between consecutive columns we need to jump forward 8 bytes in memory,
and to access the next row $3 \times 8 = 24$ bytes.
The strides of that array are therefore $(24, 8)$.

@@

NumPy's API has become ubiquitous in scientific computation, but NumPy itself
can never hope to satisfy every single specialized need of the community.
The advent of new technologies, such as GPUs, create opportunities that should be
exploited, but are out of scope for NumPy development.
Still, the widespread adoption of the NumPy API is valuable, as it lowers the
barrier to entry and provides the wider community with a stable array programming interface.
What can NumPy do to support the use of its API without blocking innovation?
We believe that having NumPy as a central point of coordination between array objects can improve interoperability while
supporting growth of a larger surrounding ecosystem of specialized solutions
and new tools.
Primarily, there exist two types of Python array objects: subclasses of NumPy
arrays, and arrays that mimic NumPy arrays but are fundamentally different.

@@

There has been a steady increase in the number of external libraries that mimic
NumPy, i.e. provide arrays and a NumPy-like API for manipulating them.
This includes popular tensor computation libraries such as
CuPy\footnote{\url{https://cupy.chainer.org/}},
JAX\footnote{\url{https://jax.readthedocs.io/en/latest/jax.numpy.html}}, and
Apache MXNet\footnote{\url{https://numpy.mxnet.io/}}.
PyTorch\footnote{\url{https://pytorch.org/tutorials/beginner/blitz/tensor\_tutorial.html}}
and
Tensorflow\footnote{\url{https://www.tensorflow.org/tutorials/customization/basics}}
provide tensor APIs with NumPy-inspired semantics.
Ideally, operating on these arrays using NumPy would simply work, so that end
users could write code once, and would then benefit from switching between
NumPy arrays, GPU arrays, distributed arrays, and so forth, as appropriate.

To facilitate \emph{interoperability}, the NumPy team is defining so-called
``protocols'' (or contracts of operation), that allow for these arrays to be
passed to NumPy functions.
NumPy, in turn, dispatches operations to the originating library, as required.  Early versions of these protocols have been successfully deployed.

%\url{https://numpy.org/neps/nep-0037-array-module.html}

>>>>

The {\em shape} of an array determines the number of elements along each axis,
and the number of axes is the array's dimensionality.
For example, a vector of numbers can be stored as a one dimensional array of
shape $N$, while color videos are four dimensional arrays of shape
$(T, M, N, 3)$.

{\em Strides} are necessary to interpret computer memory, which stores elements
linearly, as $n$-dimensional arrays.
It describes the number of bytes to move forward in memory to jump from row to
row, column to column, and so forth.
Consider, for example, a 2-D array of floating point numbers with shape
$(4, 3)$, where each element occupies 8 bytes in memory.
To move between consecutive columns we need to jump forward 8 bytes in memory,
and to access the next row $3 \times 8 = 24$ bytes.
The strides of that array are therefore $(24, 8)$.

@@

NumPy's API has become ubiquitous in scientific computation, but NumPy itself
can never hope to satisfy every single specialized need of the community.
The advent of new technologies, such as GPUs, create opportunities that should be
exploited, but are out of scope for NumPy development.

Still, the widespread adoption of the NumPy API is valuable, as it lowers the
barrier to entry and provides the wider community with a stable array programming interface.

What can NumPy do to support the use of its API without blocking innovation?
We believe that having NumPy as a central point of coordination between array objects can improve interoperability while
supporting growth of a larger surrounding ecosystem of specialized solutions
and new tools.
Primarily, there exist two types of Python array objects: subclasses of NumPy
arrays, and arrays that mimic NumPy arrays but are fundamentally different.

@@

There has been a steady increase in the number of external libraries that
provide arrays and a NumPy-like API for manipulating them.
This includes popular tensor computation libraries such as
CuPy\footnote{\url{https://cupy.chainer.org/}},
JAX\footnote{\url{https://jax.readthedocs.io/en/latest/jax.numpy.html}}, and
Apache MXNet\footnote{\url{https://numpy.mxnet.io/}}.
PyTorch\footnote{\url{https://pytorch.org/tutorials/beginner/blitz/tensor\_tutorial.html}}
and
Tensorflow\footnote{\url{https://www.tensorflow.org/tutorials/customization/basics}}
provide tensor APIs with NumPy-inspired semantics.
Ideally, operating on these arrays using NumPy would simply work, so that end
users could write code once, and would then benefit from switching between
NumPy arrays, GPU arrays, distributed arrays, and so forth, as appropriate.

To facilitate \emph{interoperability}, the NumPy team is defining so-called
``protocols'' (or contracts of operation), that allow for these arrays to be
passed to NumPy functions.
NumPy, in turn, dispatches operations to the originating library, as required.
% FIXME: expand to 3 or more sentences discussing current efforts and
% issues with those, how did adoption go, what is being done next
%\url{https://numpy.org/neps/nep-0037-array-module.html}
Early versions of these protocols have been successfully deployed.

### 2020-01-11 paper.tex
More work on outline
Update outline
Add methods section
Use array programming
Use Fig. instead of Figure in figure captions
Minor edits

\usepackage{subcaption}

@@

\title{NumPy---Array Computation for Python}

@@

\keywords{Scientific computing, Python, Mathematics}

@@

% Count of words in summary
\verbatiminput{summarycount.tex}

% Count of words in text
\verbatiminput{wordcount.tex}

@@

\section*{Data Availability}

All NumPy source code and most data generated for the current study are
available in the NumPy GitHub repository, \url{https://github.com/numpy}. Some
supporting code and data have also been stored in other public repositories
cited by this manuscript.

@@

@@

%C.R.H and T.E.O.
S.J.v.d.W and K.J.M. composed the manuscript with input from others.
S.B., K.S., T.J.R., and W.W. drafted sections.
All authors have contributed significant code, documentation, and/or expertise
to the NumPy project.
All authors reviewed the manuscript.

>>>>

\usepackage{caption}
\captionsetup{%
   figurename=Fig.,
   tablename=Tab.
}
\usepackage{subcaption}

@@

\title{NumPy---Array Programming in Python}

@@

\keywords{Scientific computing, Python, Mathematics, Array Programming}

@@

% Count of words
\verbatiminput{wordcount.tex}

@@

@@

\newpage

\input{methods}

@@

%C.R.H and T.E.O.
S.J.v.d.W and K.J.M. composed the manuscript with input from others.
S.B., K.S., T.J.R., W.W., and M.B. drafted sections.
All authors have contributed significant code, documentation, and/or expertise
to the NumPy project.
All authors reviewed the manuscript.
